\documentclass[12pt]{article}
\setlength{\parindent}{0em}
\setlength{\parskip}{0em}
\title{Roadmap}
\author{Fabian Gubler}
\date{February 2022}

\begin{document}

\maketitle

\section*{Workload}

Work week with 4 days seems most appropriate.

\begin{itemize}
	\item 12 credits amounts to 360 hours (about 40 Days). 
	\item With 4 days per week, the workload will be reached in approximately 10 weeks. 
	\item This is desirable, as it leaves me with a 2 week time buffer.\par
\end{itemize}

\section*{Part 1: Overview}
Understanding the theoretical framework of refactoring and design patterns. At the same time, get an overview of the code and relevant technologies (e.g. flask).
\begin{itemize}
	\item \textbf{Duration}: 2 Weeks
	\item \textbf{Schedule}: 1 - 3 Days in St. Gallen (depends on recommendation of Mr. Seiger)
	\item \textbf{Deliverables}: Formulation of Hypotheses, Basic understanding of Codebase
\end{itemize}


\section*{Part 2: Coding}
This is the main part of the thesis. Code refactoring based on theoretical framework. Noting down whether hypotheses are validated and documenting the results.
\begin{itemize}
	\item \textbf{Duration}: 6 Weeks
	\item \textbf{Schedule}: 3 Days in St. Gallen (preferably the entire day), 1 Day for research, tasks and writing process
	\item \textbf{Deliverables}: Refactored codebase, (Extensive) Documentation of results in the form of bullet points.
\end{itemize}

\section*{Part 3: Writing Process}
Formulation of notes into a first draft, which will then be revised and improved upon.
\begin{itemize}
	\item \textbf{Duration}: 2 Weeks \
	\item \textbf{Schedule}: Entirely remote\
	\item \textbf{Deliverables}: Final Paper\
\end{itemize}

\end{document}
